---
title: "QFish Assignment"
author: "Michael Guijt"
date: "2025-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# QFish Plotting Assignment

## Data Location

The data was sourced from the QFish government website; https://qfish.fisheries.qld.gov.au/

It is found in the subset "Commercial Harvest Fisheries," filtered to include the licence and catch data for the freshwater eel fishery.

## Data Import

First step is to import the data.

```{r}
eels <- read.csv("export.csv")
```

## Data Wrangling

The Licences and Tonnes variables that we are interested in are listed as character vectors due to "N/A." Consequently, we change all N/As with zeroes to convert the data to a numerical vector.

```{r}
#converting N/A to 0
eels <- eels %>%
  mutate(
    Tonnes = as.numeric(replace(Tonnes, Tonnes == "N/A", 0)),
    Licences = as.numeric(replace(Licences, Licences == "N/A",0))
  )
```

Next, we see that the time series is broken due to a missing value in 1998. As it is just one value, we will replace the data with the averages of the values in 1997 and 1999. 

```{r}
#replacing 1998 value with average
eels <- eels %>%
  mutate(
    Tonnes = ifelse(CalendarYear == 1998, (Tonnes[CalendarYear == 1997] + Tonnes[CalendarYear == 1999]) / 2, Tonnes),
    Licences = ifelse(CalendarYear == 1998, (Licences[CalendarYear == 1997] + Licences[CalendarYear == 1999]) / 2, Licences)
  )
```

As the time series has no data from 2019 onwards, we remove those years from the dataset.

```{r}
#only keeps values of years before 2019
eels <- eels %>%
  filter(CalendarYear < 2019)
```

Finally, we convert Year from a character to a numerical variable.

```{r}
eels$CalendarYear <- as.numeric(eels$CalendarYear)
```


## Plotting the data

Now we can plot the data using ggplot.

```{r}
ggplot(eels, aes(x=CalendarYear)) +
  geom_line(aes(y=Tonnes, color = "Tonnes"), size = 1.2) +
  geom_line(aes(y=Licences, color = "Licences"), size = 1.2, linetype = "dashed") +
  scale_y_continuous(name = "Catch (tonnes)", sec.axis = sec_axis(~ ., name = "Number of licences")) +
  labs(title = "Freshwater eel catch and number of licences over time", x = "Year") +
  scale_color_manual(values = c("Tonnes" = "darkblue", "Licences" = "darkred")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_blank(),
        legend.position = "top")
```