---
title: "QFish Assignment"
author: "Michael Guijt"
date: "2025-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# QFish Plotting Assignment

First step is to import the data.

```{r}
eels <- read.csv("export.csv")
```

The Licences and Tonnes variables that we are interested in are listed as character vectors due to "N/A." Consequently, we change all N/As with zeroes to convert the data to a numerical vector.

```{r}
#converting N/A to 0
eels <- eels %>%
  mutate(
    Tonnes = as.numeric(replace(Tonnes, Tonnes == "N/A", 0)),
    Licences = as.numeric(replace(Licences, Licences == "N/A",0))
  )
```

Next, we see that the time series is broken due to a missing value in 1998. As it is just one value, we will replace the data with the averages of the values in 1997 and 1999. 

```{r}
#replacing 1998 value with average
eels <- eels %>%
  mutate(
    Tonnes = ifelse(CalendarYear == 1998, (Tonnes[CalendarYear == 1997] + Tonnes[CalendarYear == 1999]) / 2, Tonnes),
    Licences = ifelse(CalendarYear == 1998, (Licences[CalendarYear == 1997] + Licences[CalendarYear == 1999]) / 2, Licences)
  )
```