---
title: "MB5370 Module 04. Workshop 2 - Using ggplot2 for communication"
author: "Michael Guijt"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.1 Labels

Adding labels;

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
```

Adding subtitle and caption.

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size", 
       subtitle = "Two seaters (sports cars) are an exception because of their light weight",
       caption = "Data from fueleconomy.gov")
```

Using the labs() function to replace axis labels and legend titles.

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```

## 3.2 Annotations

Adding text annotations of a specified data value to the plot itself.

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)
```

## 3.3 Scales

Showing that scales are automatically added.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
```

Adding custom axis limits;

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(limits = c(0,12)) +
  scale_y_continuous() +
  scale_colour_discrete()
```

## 3.4 Axis ticks

Using breaks to change the position of axis markers.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
```

Can also set labels to NULL to remove them.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```

## 3.5 Legends and colour schemes

Using theme() to change the position of the legend.

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) 

#trying different legend positions
base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right") #the default
```

## 3.6 Replacing the scale

Demonstration on transforming scale; converting the carat and price data of the diamonds to log scale;

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() +
  scale_x_log10() +
  scale_y_log10()
```

Customising colour scales;

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_colour_brewer(palette = "Set1")
```

Adding shapes can make your graph more interpretable for colorblind people;

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set1")
```

Setting manual colors without calling in a palette;

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

Using the viridis colour package, very popular;

```{r}
#making a fake dataset
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x,y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x,y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```

## 3.7 Themes

Showcasing different themes:

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_bw()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_light()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_classic()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_dark()
```

Example of determining your own theme;

```{r, message=FALSE}
theme (panel.border = element_blank(),
       panel.grid.minor.x = element_blank(),
       panel.grid.minor.y = element_blank(),
       legend.position = "bottom",
       legend.title = element_blank(),
       legend.text = element_text(size=8),
       panel.grid.major = element_blank(),
       axis.text.y = element_text(colour="black"),
       axis.text.x = element_text(colour="black"),
       text = element_text(family="Arial"))
```

r
