---
title: "MB5370 Module 04. Workshop 2 - Using ggplot2 for communication"
author: "Michael Guijt"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.1 Labels

Adding labels;

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
```

Adding subtitle and caption.

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size", 
       subtitle = "Two seaters (sports cars) are an exception because of their light weight",
       caption = "Data from fueleconomy.gov")
```

Using the labs() function to replace axis labels and legend titles.

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```

## 3.2 Annotations

Adding text annotations of a specified data value to the plot itself.

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)
```

## 3.3 Scales

Showing that scales are automatically added.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
```

Adding custom axis limits;

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous(limits = c(0,12)) +
  scale_y_continuous() +
  scale_colour_discrete()
```

## 3.4 Axis ticks

Using breaks to change the position of axis markers.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
```

Can also set labels to NULL to remove them.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```

## 3.5 Legends and colour schemes

